16s beta diversity

### 2021-03-08 scripts/45_16s_sorting_out.Rmd
ok slowly picking up the work

uptb <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/U Penn/patients/upenn_essential_data_v3.xlsx', sheet = 3) %>% 
  filter(!is.na(gender))

@@

```{r}
m <- m_meta_b %>% 
  select(-pt_id_old) %>% 
  rename(pt_id = mrn) %>% 
  select(-last_followup) %>% 
  mutate(center = 'M') %>% 
  mutate(icansgrade = as.character(icansgrade))

@@

```{r}
get_table_from_database('asv_alpha_diversity_ag')

>>>>

uptb <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/U Penn/patients/upenn_essential_data_v3.xlsx', sheet = 3) %>% 
  filter(sequenced == 'yes')
```

@@

```{r}
m <- m_meta_b %>% 
  select(-pt_id_old) %>% 
  rename(pt_id = mrn) %>% 
  select(-last_followup) %>% 
  mutate(center = 'M') %>% 
  select(-cr_overall) # this column is now gone

@@

```{r}
library(vdbR)
connect_database('~/dbConfig.txt')
get_table_from_database('asv_alpha_diversity_ag')

### 2021-03-08 scripts/48_16s_beta_diversity.Rmd
ok slowly picking up the work

dat <- genus %>% 
  filter(genus %in% keep) %>% 
  spread(key = 'genus', value = 'relab') %>% 
  full_join(pheno)

@@

```{r}
# a function to do the combinations of the anosim test
pheno4 <- colnames(pheno)[2:5]

combi <- data_frame(
  g1 = c(rep('M',4), rep('P',4), rep('M',4)),
  g2 = c(rep('M',4), rep('P',4), rep('P',4)),
  label = rep(pheno4, 3)
)

do_anosim_each <-  function(g1_, g2_, label_){
    df = dat %>% 
      filter(center %in% c(g1_, g2_)) %>% 
      filter(get(label_) %in% c('yes','no')) 
    cts = df %>% 
      select(sampleid, all_of(keep)) %>% 
      column_to_rownames('sampleid')
    phe = df %>% 
      select(matches(label_)) %>% pull(get(label_)) 
    phedf = df %>% 
      select(sampleid, matches(label_))

    # anosim run
    ret = anosim(cts, phe, permutations = 999, distance = "bray", strata = NULL, parallel = 4)

    # plot the pcoa
    dist_ <- vegdist(cts, method = 'bray')
    bc <- cmdscale(dist_, k = 2) 
    bc %>% 
      as.data.frame() %>% 
      rownames_to_column('sampleid') %>% 
      full_join(phedf, by = 'sampleid') %>% 
      ggplot(aes_string(x = "V1", y = "V2", color =  label_)) + geom_point() +
      labs(title = str_glue('PCoA {g1_}_{g2_}__{label_}\npval:{ret$signif} statistic:{round(ret$statistic, 2)}')) +
      ggsave(str_glue('../figs/beta_{g1_}_{g2_}__{label_}.pdf'))

  return(list(pval = ret$signif, 
              statistic = ret$statistic))
}

@@

>>>>

# find the dominating genus in each sample
domi <- genus %>% 
  arrange(sampleid, -relab) %>% 
  group_by(sampleid) %>% 
  slice(1) %>% 
  rename(domgenus = genus) %>% 
  select(sampleid, domgenus)

dat <- genus %>% 
  filter(genus %in% keep) %>% 
  spread(key = 'genus', value = 'relab') %>% 
  full_join(pheno) %>% 
  full_join(domi)

@@

```{r}
# a function to do the combinations of the anosim test
pheno_ <- c(colnames(pheno)[2:5] , 'domgenus')
n <- length(pheno_)

combi <- data_frame(
  g1 = c(rep('M',n), rep('P',n), rep('M',n)),
  g2 = c(rep('M',n), rep('P',n), rep('P',n)),
  label = rep(pheno_, 3)
)

do_anosim_each <-  function(g1_, g2_, label_){
  g1_ = 'M'
  g2_ = 'P'
  label_ = 'domgenus'
    df = dat %>% 
      filter(center %in% c(g1_, g2_)) %>% 
      filter(get(label_) %in% c('yes','no')) 
    cts = df %>% 
      select(sampleid, all_of(keep)) %>% 
      column_to_rownames('sampleid')
    phe = df %>% 
      select(matches(label_)) %>% pull(get(label_)) 
    phedf = df %>% 
      select(sampleid, matches(label_))
    # 
    # # anosim run
    # ret = anosim(cts, phe, permutations = 999, distance = "bray", strata = NULL, parallel = 4)
    # 
    # # plot the pcoa
    dist_ <- vegdist(cts, method = 'bray')
    bc <- cmdscale(dist_, k = 2)
    # bc %>% 
    #   as.data.frame() %>% 
    #   rownames_to_column('sampleid') %>% 
    #   full_join(phedf, by = 'sampleid') %>% 
    #   ggplot(aes_string(x = "V1", y = "V2", color =  label_)) + geom_point() +
    #   labs(title = str_glue('PCoA {g1_}_{g2_}__{label_}\npval:{ret$signif} statistic:{round(ret$statistic, 2)}')) +
    #   ggsave(str_glue('../figs/beta_{g1_}_{g2_}__{label_}.pdf'))

    # the version without the anosim stats
    bc %>% 
      as.data.frame() %>% 
      rownames_to_column('sampleid') %>% 
      full_join(phedf, by = 'sampleid') %>% 
      ggplot(aes_string(x = "V1", y = "V2", color =  label_)) + geom_point() +
      labs(title = str_glue('PCoA {g1_}_{g2_}__{label_}')) +
      ggsave(str_glue('../figs/beta_{g1_}_{g2_}__{label_}.pdf'))

  # return(list(pval = ret$signif, 
  #             statistic = ret$statistic))
}

@@

```{r}
# do the pcoa colored by the dominating genus 
do_domgenus_coloring_each <-  function(g1_, g2_, label_){
  g1_ = 'M'
  g2_ = 'P'
  label_ = 'domgenus'
    df = dat %>% 
      filter(center %in% c(g1_, g2_)) 
    cts = df %>% 
      select(sampleid, all_of(keep)) %>% 
      column_to_rownames('sampleid')
    phe = df %>% 
      select(matches(label_)) %>% pull(get(label_)) 
    phedf = df %>% 
      select(sampleid, matches(label_))

    # # plot the pcoa
    dist_ <- vegdist(cts, method = 'bray')
    bc <- cmdscale(dist_, k = 2)

    # the version without the anosim stats
    bc %>% 
      as.data.frame() %>% 
      rownames_to_column('sampleid') %>% 
      full_join(phedf, by = 'sampleid') %>% 
      ggplot(aes_string(x = "V1", y = "V2", color =  label_)) + geom_point() +
      labs(title = str_glue('PCoA  combined {label_}')) +
      ggsave(str_glue('../figs/beta_{g1_}_{g2_}__{label_}.pdf'))

}
```
